---
title: "longRNA_Oasis_Clustering"
author: "Michael"
date: "15 1 2020"
output:
    html_document:
        toc: true
        toc_float: true
        theme: 'cerulean'
        highlight: 'tango'
        code_folding: hide
        df_print: paged
---


```{r}
suppressPackageStartupMessages({
  library(edgeR)
  library(SummarizedExperiment)
  library(SEtools)
  library(ggplot2)
  library(matrixStats)
  library(igraph)
  library(RUVSeq)
  library(viridis)
  library(RColorBrewer)
  library(data.table)
  library(dplyr)
  library(tibble)
  library(pheatmap)
  library(plgINS)
  library(sechm)
  library(stringr)
  library(ComplexHeatmap)
})
source("../../functions/clustering.R")
source("../../functions/clusterEnrichment.R")
```



# Color Definitions
```{r}
ancols <- list(day=c("0"="#3F3F3F",
                       "9" ="#575757",
                       "15" ="#707070",
                       "21" = "#898989",
                       "27" ="#A1A1A1",
                       "33" ="#BABABA",
                       "40" = "#D3D3D3"))


ancols2 <- list(day=c("0"="#3F3F3F",
                       "9" ="#575757",
                       "15" ="#707070",
                       "21" = "#898989",
                       "27" ="#A1A1A1",
                       "33" ="#BABABA",
                       "40" = "#D3D3D3"),
               cluster=c("1" = "#440154FF",
                         "2" = "#482878FF",
                         "3" = "#3E4A89FF",
                         "4" = "#31688EFF",
                         "5" = "#26828EFF",
                         "6" = "#1F9E89FF",
                         "7" = "#35B779FF",
                         "8" = "#6DCD59FF",
                         "9" = "#B4DE2CFF",
                         "10" = "#FDE725FF"
                         ))



heatcol = inferno(255)
```



## Import datasets

```{r}
set.seed(1234)
se <- readRDS("../data/Long_RNA/longRNA_noRound.DEA.SE.rds")
se$day <- as.factor(se$day)
se2 <- se[,se$day != 0]
```

Significants
```{r}
# select significant results
## all stages
res.all <- rowData(se)$DEA.all
res.all$maxlfc <- rowMax(abs(as.matrix(res.all[,1:6]))) 
sig.all <- row.names(res.all)[res.all$FDR < 0.01 & res.all$maxlfc > 2]
sig.all_plot <- row.names(res.all)[res.all$FDR < 0.01 & res.all$maxlfc > 2 & res.all$logCPM > 3]

## neuronal stages
res.neu <- rowData(se)$DEA.neuronal
res.neu$maxlfc <- rowMax(abs(as.matrix(res.neu[,1:5]))) 
sig.neu <- row.names(res.neu)[res.neu$FDR < 0.01 & res.neu$maxlfc > 2]
```

# Clustering all days

## Get preliminary Cluster Info
# Pay attention, takes a very long time
```{r}
## prepare logFC matrix
lfc <- prepLFC4clust(assays(se)$log2FC.all[sig.all,], groups=se$stage, minLFC = 0.5)
## clustering & analysis
png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/Cluster_Info_all-days.png", width=7, height=5, units="in", res=300)
cc <- clusterWrapper(lfc, 2:15, "pam")
dev.off()
```


## Get clusters and save object
```{r}
rowData(se)$cluster <- as.factor(cc[["11"]]$cluster[row.names(se)])
saveRDS(se,"../data/Long_RNA/longRNA_noRound.DEA.SE.cluster.rds")
```



# Load se object
```{r}
se <- readRDS("../data/Long_RNA/longRNA_noRound.DEA.SE.cluster.rds")
```

## smooth plots
```{r}
sm <- getSmoothedMeans(assays(se)$logcpm[sig.all,], as.numeric(factor(se$day)), method = "loess", f=0.8)
plotClusterLFCs(sm-rowMeans(sm), rowData(se[sig.all,])$cluster)

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_all-days.png", width=7, height=5, units="in", res=300)
plotClusterLFCs(sm-rowMeans(sm), rowData(se[sig.all,])$cluster)
dev.off()
```

# Cluster Enrichment Analysis
```{r}
neu.all <- unique(gsub(".*\\.", "", rownames(se)))
all.clus <- split(sapply(strsplit(row.names(se), "\\."), FUN=function(x) rev(x)[1]),rowData(se)$cluster)
all.clus.un <- lapply(all.clus, unique)
gsets <- plgINS::getMsigSets(collections=c("C5"))
gsets_cl <- gsets[!grepl("C5:HPO",names(gsets))]
names(gsets_cl) <- gsub("C5:GO:","",names(gsets_cl))
dgo.all <- clusterEnrichment(all.clus.un, gsets_cl)
```

```{r}
png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_ClusterEnrichment_all-days.png", width=12, height=7, units="in", res=300)
plotClusterEnrichment(dgo.all,color = heatcol)
dev.off()
```

# lncRNAs from Cluster9
```{r}
cl2.all.ncRNA <- row.names(se)[which(rowData(se)$cluster=="2" & rowData(se)$biotype=="lincRNA")]

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_All_Cluster2_lincRNA_all-days.png", width=7, height=5, units="in", res=300)
sehm(se, cl2.all.ncRNA , assayName = "logcpm", do.scale=TRUE, anno_columns = c("day"), anno_rows = c("logcpm"), breaks=0.99, main="All Cluster 2 lncRNAs",hmcols = heatcol, anno_colors = ancols,annotation_names_col = FALSE)
dev.off()
```





# Clustering neuronal days object

## Get preliminary Cluster Info
```{r}
## prepare logFC matrix
lfc2 <- prepLFC4clust(assays(se2)$log2FC.neur[sig.neu,], groups=se2$stage, minLFC = 0.5)
lfc2 <- lfc2[,-1]
## clustering & analysis
png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/Cluster_Info_Neur-days.png", width=7, height=5, units="in", res=300)
cc2 <- clusterWrapper(lfc2, 2:15, "pam")
dev.off()
```

## Get clusters and save object
```{r}
rowData(se2)$cluster <- as.factor(cc2[["10"]]$cluster[row.names(se2)])
saveRDS(se2,"../data/Long_RNA/longRNA_noRound.DEA.SE2.neur_days.cluster.rds")
```

# Load se2 object
```{r}
se2 <- readRDS("../data/Long_RNA/longRNA_noRound.DEA.SE2.neur_days.cluster.rds")
```


## smooth plots
```{r}
sm2 <- getSmoothedMeans(assays(se2)$logcpm[sig.neu,], as.numeric(factor(se2$day)), method = "loess", f=0.8)
plotClusterLFCs(sm2-rowMeans(sm2), rowData(se2[sig.neu,])$cluster)

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_neur-days.png", width=7, height=5, units="in", res=300)
plotClusterLFCs(sm2-rowMeans(sm2), rowData(se2[sig.neu,])$cluster)
dev.off()
```



## Heatmap - Plot all neuronal Clusters
```{r, fig.height=10, fig.width=6}
sechm(se2, sig.neu, assayName = "logcpm", do.scale=TRUE, top_annotation = c("day"),annocol_title_side=NA, right_annotation = c("cluster"),  breaks=0.99, toporder = "cluster",hmcols = heatcol, anno_colors = ancols2)

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_Heat_neur-days.png", width=7, height=7, units="in", res=300)
sechm(se2, sig.neu, assayName = "logcpm", do.scale=TRUE, top_annotation = c("day"),annocol_title_side=NA, right_annotation = c("cluster"),  breaks=0.99, toporder = "cluster",hmcols = heatcol, anno_colors = ancols2)
dev.off()
```


# Cluster Enrichment Analysis
```{r}
neu.un <- unique(gsub(".*\\.", "", rownames(se2)))
neu.clus <- split(sapply(strsplit(row.names(se2), "\\."), FUN=function(x) rev(x)[1]),rowData(se2)$cluster)
neu.clus.un <- lapply(neu.clus, unique)
gsets <- plgINS::getMsigSets(collections=c("C5"))
gsets_cl <- gsets[!grepl("C5:HPO",names(gsets))]
names(gsets_cl) <- gsub("C5:GO:","",names(gsets_cl))
dgo.neu <- clusterEnrichment(neu.clus.un, gsets_cl)
```

```{r}
png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_ClusterEnrichment_neur-days.png", width=12, height=7, units="in", res=300)
plotClusterEnrichment(dgo.neu,color = heatcol)
dev.off()
```

```{r}
#Plot a clean version
dgo.neu2 <- dgo.neu
row.names(dgo.neu2) <- gsub("GO_","",row.names(dgo.neu2))
row.names(dgo.neu2) <- gsub("GOBP_","",row.names(dgo.neu2))
row.names(dgo.neu2) <- gsub("GOCC_","",row.names(dgo.neu2))
row.names(dgo.neu2) <- gsub("GOMF_","",row.names(dgo.neu2))

a <- plotClusterEnrichment(dgo.neu2,color = heatcol,trim.names = 35L,row_names_max_width = unit(9, "cm"),)

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_ClusterEnrichment_neur-days_names_clean_CompHeat.png", width=7, height=6, units="in", res=300)
  draw(a,heatmap_legend_side="left")
dev.off()
```







## Get the Go-Terms

# Single Go
```{r}
ccg.neu <- split(row.names(se2),rowData(se2)$cluster)
c9.neu <- unique(as.character(rowData(se2)[ccg.neu[["9"]],"symbol"]))
allg <- unique(as.character(rowData(se2)$symbol))
allg <- allg[!is.na(allg)]
go9 <- plgINS::goseq.enrichment(allg, c9.neu, gotype = "GO:CC")
head(go9)
```



# Single Go
```{r}
ccg.neu <- split(row.names(se2),rowData(se2)$cluster)
c6.neu <- unique(as.character(rowData(se2)[ccg.neu[["6"]],"symbol"]))
allg <- unique(as.character(rowData(se2)$symbol))
allg <- allg[!is.na(allg)]
go6 <- plgINS::goseq.enrichment(allg, c6.neu, gotype = "GO:CC")
head(go6)
```


# All GO List
```{r}
ccg.neu <- split(row.names(se2),rowData(se2)$cluster)
allg <- unique(as.character(rowData(se2)$symbol))
allg <- allg[!is.na(allg)]

GO.neu.list <- list()
for(i in 1:10){
  c.neu <- unique(as.character(rowData(se2)[ccg.neu[[i]],"symbol"]))
  go <- plgINS::goseq.enrichment(allg, c.neu, gotype = "GO:CC")
  GO.neu.list[[i]] <- go
}

saveRDS(GO.neu.list,"../data/Long_RNA/GO_Terms/GO_Terms_NeuronalCluster_longRNA.rds")
```




# Plot single neuronal Cluster
```{r}
cluster.neu <- split(sig.neu, rowData(se2[sig.neu,])$cluster)
```


# Cluster 9 
```{r, fig.height=8, fig.width=6}
cluster.neu.9 <- cluster.neu[["9"]]
se_cl9 <- se2[row.names(se2) %in% cluster.neu.9,]
saveRDS(se_cl9,"../data/Long_RNA/longRNA_noRound.DEA.SE_cluster9.rds")

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_Neuronal_Cluster9_neur-days.png", width=7, height=5, units="in", res=300)
sehm(se2, cluster.neu.9 , assayName = "logcpm", do.scale=TRUE, anno_columns = c("day"), anno_rows = c("logcpm"), breaks=0.99, main="Neuronal Cluster 9",hmcols = heatcol, anno_colors = ancols,annotation_names_col = FALSE)
dev.off()

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_Neuronal_Cluster9_all-days.png", width=7, height=5, units="in", res=300)
sehm(se, cluster.neu.9 , assayName = "logcpm", do.scale=TRUE, anno_columns = c("day"), anno_rows = c("logcpm"), breaks=0.99, main="Neuronal Cluster 9",hmcols = heatcol, anno_colors = ancols,annotation_names_col = FALSE)
dev.off()
```

# lncRNAs from Cluster9
```{r}
cl9.ncRNA <- row.names(se2)[which(rowData(se2)$cluster=="9" & rowData(se2)$biotype=="lincRNA")]

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_Neuronal_Cluster9_lincRNA_all-days.png", width=7, height=5, units="in", res=300)
sehm(se, cl9.ncRNA , assayName = "logcpm", do.scale=TRUE, anno_columns = c("day"), anno_rows = c("logcpm"), breaks=0.99, main="Neuronal Cluster 9 lncRNAs",hmcols = heatcol, anno_colors = ancols,annotation_names_col = FALSE)
dev.off()
```


# lncRNAs from Cluster9, maxlfc > 3
```{r}
cl9.ncRNA <- row.names(se2)[which(rowData(se2)$cluster=="9" & rowData(se2)$biotype=="lincRNA")]
cl9.ncRNA.maxLFC <- intersect(cl9.ncRNA,row.names(res.neu)[res.neu$maxlfc > 4])

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_Neuronal_Cluster9_lincRNA_all-days_maxlfc4.png", width=7, height=5, units="in", res=300)
sehm(se, cl9.ncRNA.maxLFC , assayName = "logcpm", do.scale=TRUE, anno_columns = c("day"), anno_rows = c("logcpm"), breaks=0.99, main="Neuronal Cluster 9 lncRNAs (maxlFC > 4)",hmcols = heatcol, anno_colors = ancols,annotation_names_col = FALSE)
dev.off()
```


# lncRNAs from Cluster7
```{r}
cl7.ncRNA <- row.names(se2)[which(rowData(se2)$cluster=="7" & rowData(se2)$biotype=="lincRNA")]

png("../Figures_Omics/Long_RNAs/Figure_Output/Cluster/DEG_PAM_Neuronal_Cluster7_lincRNA_all-days.png", width=7, height=5, units="in", res=300)
sehm(se, cl7.ncRNA , assayName = "logcpm", do.scale=TRUE, anno_columns = c("day"), anno_rows = c("logcpm"), breaks=0.99, main="Neuronal Cluster 7 lncRNAs",hmcols = heatcol, anno_colors = ancols,annotation_names_col = FALSE)
dev.off()
```




