---
title: "Proteomics_LongRNA_Corr"
author: "Michael Soutschek"
date: "11/23/2022"
output: html_document
---




```{r}
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(SEtools)
  library(tidyverse)
  library(ggpubr)
  library(ggplot2)
  library(plgINS)
  library(viridis)
  library(cowplot)
  library(ComplexHeatmap)
  library(sechm)
})
```

# Color Definitions
```{r}
ancols <- list(Day=c("0"="#3F3F3F",
                       "9" ="#575757",
                       "15" ="#707070",
                       "21" = "#898989",
                       "27" ="#A1A1A1",
                       "33" ="#BABABA",
                       "40" = "#D3D3D3" 
) )

ancols2 <- list(Day=c(
                       "9" ="#575757",
                       "15" ="#707070",
                       "21" = "#898989",
                       "27" ="#A1A1A1",
                       "33" ="#BABABA",
                       "40" = "#D3D3D3" 
) )


heatcol = inferno(255)
```

## Load data
```{r}
se_RNA <- readRDS("../../RNA_GRCh38.p10_Ens91/data/Long_RNA/longRNA.DEA.SE.rds")
se_RNA$day <- as.factor(se_RNA$day)
se_PROT <- readRDS("../../Proteomics/data/Proteomics/Proteomics.Reference.SE_DEA.rds")
se_PROT$day <- as.factor(se_PROT$day)
```


```{r significant results, eval=TRUE}
# select significant results
## all stages RNA
res <- rowData(se_RNA)$DEA.all
res$maxlfc <- rowMax(abs(as.matrix(res[,1:6]))) 
sig_all <- row.names(res)[res$FDR < 0.01]
sig_all <- sapply(strsplit(sig_all,"\\."),"[",2)

sig_all_lFC <- row.names(res)[res$FDR < 0.01 & res$maxlfc > 2]
sig_all_lFC <- sapply(strsplit(sig_all_lFC,"\\."),"[",2)

## neuronal stages
res.neu <- rowData(se_RNA)$DEA.neuronal
res.neu$maxlfc <- rowMax(abs(as.matrix(res.neu[,1:5]))) 
sig.neu <- row.names(res.neu)[res.neu$FDR < 0.01]
sig.neu <- sapply(strsplit(sig.neu,"\\."),"[",2)

sig.neu_lFC <- row.names(res.neu)[res.neu$FDR < 0.01 & res.neu$maxlfc > 2]
sig.neu_lFC <- sapply(strsplit(sig.neu_lFC,"\\."),"[",2)
```



## Prepare SEs for RNA and Protein to make them comparable 
```{r}
## sort dataframe columns of rowdata as they may be in different order from the SE itself
sortSE <- function(se) {
  lapply(names(rowData(se)), function(colname) {
    col <- rowData(se)[[colname]]
    if(class(col) %in% c("data.frame", "DFrame")) rowData(se)[[colname]] <<- col[names(se), ]
  })
  return(se)
}

se_RNA <- sortSE(se_RNA)
se_PROT <- sortSE(se_PROT)


## align coldata
colnames(se_PROT) <- colnames(se_RNA)
cd_merged <- merge(colData(se_PROT), colData(se_RNA), by = c("row.names", "day", "stage"), sort = FALSE)
row.names(cd_merged) <- cd_merged$Row.names
cd_merged$Row.names <- NULL
colData(se_PROT) <- cd_merged[names(cd_merged)!="libsize"]
colData(se_RNA) <- cd_merged

## sort columns by stage
se_RNA <- se_RNA[, order(se_RNA$day)]
se_PROT <- se_PROT[, order(se_PROT$day)]

## reduce to common genes
common_genes <- intersect(rowData(se_RNA)$symbol, rowData(se_PROT)$GeneName)
se_RNA_red <- se_RNA[rowData(se_RNA)$symbol %in% common_genes,]
se_PROT_red <- se_PROT[rowData(se_PROT)$GeneName %in% common_genes,]

## remove duplicate genes 
dubl_PROT <- rowData(se_PROT_red)$GeneName[duplicated(rowData(se_PROT_red)$GeneName)]
dubl_RNA <- rowData(se_RNA_red)$symbol[duplicated(rowData(se_RNA_red)$symbol)]
ambiguous_entries <- c(dubl_PROT,dubl_RNA)

dubl_row_PROT <- row.names(rowData(se_PROT_red))[rowData(se_PROT_red)$GeneName %in% ambiguous_entries]
dubl_row_RNA <- row.names(rowData(se_RNA_red))[rowData(se_RNA_red)$symbol %in% ambiguous_entries]

se_PROT_red <- se_PROT_red[!(row.names(se_PROT_red) %in% dubl_row_PROT)]
se_RNA_red <- se_RNA_red[!(row.names(se_RNA_red) %in% dubl_row_RNA)]

## aggregate SEs by gene names
se_RNA_agg <- aggSE(se_RNA_red, by = "symbol")
se_PROT_agg <- aggSE(se_PROT_red, by = "GeneName")
assays(se_RNA_agg)[["expression"]] <- assays(se_RNA_agg)[["logcpm"]]
assays(se_PROT_agg)[["expression"]] <- assays(se_PROT_agg)[["imputed"]]

colnames(colData(se_RNA_agg))[which(colnames(colData(se_RNA_agg)) == "day")] <- "Day"
colnames(colData(se_PROT_agg))[which(colnames(colData(se_PROT_agg)) == "day")] <- "Day"
```


#Prepare Data.Frames to plot
```{r}
RNA_long <- assays(se_RNA_agg)[["expression"]] %>%
  as.data.frame() %>%
  rownames_to_column("Gene") %>%
  pivot_longer(-Gene, names_to = "condition", values_to = "RNA")

PROT_long <- assays(se_PROT_agg)[["expression"]] %>%
  as.data.frame() %>%
  rownames_to_column("Gene") %>%
  pivot_longer(-Gene, names_to = "condition", values_to = "PROT")

RNA_PROT_merged <- merge(RNA_long, PROT_long, by = c("Gene", "condition")) %>%
  mutate(Day = as.factor(as.numeric(gsub(".*_", "", condition)))) %>%
  mutate(replicate = as.factor(gsub("_.*", "", condition)))

RNA_PROT_merged.neu <- RNA_PROT_merged[RNA_PROT_merged$Day != 0,]

RNA_PROT_merged_long <- RNA_PROT_merged %>%
  pivot_longer(RNA:PROT, names_to = "type", values_to = "expression")

RNA_PROT_merged_long.neu <- RNA_PROT_merged_long[RNA_PROT_merged_long$Day != 0,]
```


#Prepare Plotting dataframes for all Days
```{r}
RNA_PROT_mean <- RNA_PROT_merged_long %>%
  group_by(Gene, Day, type) %>%
  summarize(mean_exp = mean(expression)) %>%
  group_by(Gene, type) %>%
  mutate(var = var(mean_exp)) %>%
  mutate(norm_exp = scale(mean_exp))

# RNA_PROT_cor <- RNA_PROT_merged %>%
#   group_by(Gene, Day) %>%
#   summarize(RNA_mean = mean(RNA), PROT_mean = mean(PROT)) %>%
#   group_by(Gene) %>%
#   summarize(pearson = cor(RNA_mean, PROT_mean, method = "pearson"),
#             spearman = cor(RNA_mean, PROT_mean, method = "spearman"),
#             var_RNA = var(RNA_mean), var_PROT = var(PROT_mean))
# 
# RNA_PROT_cor_fil <- RNA_PROT_merged %>%
#   group_by(Gene, Day) %>%
#   summarize(RNA_mean = mean(RNA), PROT_mean = mean(PROT)) %>%
#   group_by(Gene) %>%
#   summarize(pearson = cor(RNA_mean, PROT_mean, method = "pearson"),
#             spearman = cor(RNA_mean, PROT_mean, method = "spearman"),
#             var_RNA = var(RNA_mean), var_PROT = var(PROT_mean)) %>%
#   dplyr::filter(var_RNA > 0.1 & var_PROT > 0.1)
# 
# RNA_PROT_cor.sig <- RNA_PROT_cor[RNA_PROT_cor$Gene %in% sig_all,]
# RNA_PROT_cor.sig_lFC <- RNA_PROT_cor[RNA_PROT_cor$Gene %in% sig_all_lFC,]
# 
# #is basically the same as the filtered
# #RNA_PROT_cor.fil.sig <- RNA_PROT_cor_fil[RNA_PROT_cor_fil$Gene %in% sig_all,]
# 
# most_uncorrelated <- RNA_PROT_cor %>% arrange(pearson) %>% pull(Gene) %>% head(12)
# most_uncorrelated.sig <- RNA_PROT_cor.sig %>% arrange(pearson) %>% pull(Gene) %>% head(12)
# most_uncorrelated.sig_lFC <- RNA_PROT_cor.sig_lFC %>% arrange(pearson) %>% pull(Gene) %>% head(12)
# most_uncorrelated.fil <- RNA_PROT_cor_fil %>% arrange(pearson) %>% pull(Gene) %>% head(12)
# 
# most_correlated <- RNA_PROT_cor %>% arrange(desc(pearson)) %>% pull(Gene) %>% head(12)
```



#Prepare Plotting dataframes for neuronal Days
```{r}
RNA_PROT_mean.neu <- RNA_PROT_merged_long.neu %>%
  group_by(Gene, Day, type) %>%
  summarize(mean_exp = mean(expression)) %>%
  group_by(Gene, type) %>%
  mutate(var = var(mean_exp)) %>%
  mutate(norm_exp = scale(mean_exp))

RNA_PROT_cor.neu <- RNA_PROT_merged.neu %>%
  group_by(Gene, Day) %>%
  summarize(RNA_mean = mean(RNA), PROT_mean = mean(PROT)) %>%
  group_by(Gene) %>%
  summarize(pearson = cor(RNA_mean, PROT_mean, method = "pearson"),
            spearman = cor(RNA_mean, PROT_mean, method = "spearman"),
            var_RNA = var(RNA_mean), var_PROT = var(PROT_mean))

RNA_PROT_cor_fil.neu <- RNA_PROT_merged.neu %>%
  group_by(Gene, Day) %>%
  summarize(RNA_mean = mean(RNA), PROT_mean = mean(PROT)) %>%
  group_by(Gene) %>%
  summarize(pearson = cor(RNA_mean, PROT_mean, method = "pearson"),
            spearman = cor(RNA_mean, PROT_mean, method = "spearman"),
            var_RNA = var(RNA_mean), var_PROT = var(PROT_mean)) %>%
  dplyr::filter(var_RNA > 0.1 & var_PROT > 0.1)

RNA_PROT_cor.neu.sig <- RNA_PROT_cor.neu[RNA_PROT_cor.neu$Gene %in% sig.neu,]
RNA_PROT_cor.neu.sig_lFC <- RNA_PROT_cor.neu[RNA_PROT_cor.neu$Gene %in% sig.neu_lFC,]

#most_uncorrelated.neu <- RNA_PROT_cor.neu %>% arrange(pearson) %>% pull(Gene) %>% head(9)
most_uncorrelated.neu.sig <- RNA_PROT_cor.neu.sig %>% arrange(pearson) %>% pull(Gene) %>% head(8)
most_uncorrelated.neu.sig_lFC <- RNA_PROT_cor.neu.sig_lFC %>% arrange(pearson) %>% pull(Gene) %>% head(9)
most_uncorrelated.fil.neu <- RNA_PROT_cor_fil.neu %>% arrange(pearson) %>% pull(Gene) %>% head(9)

most_correlated.neu <- RNA_PROT_cor.neu %>% arrange(desc(pearson)) %>% pull(Gene) %>% head(12)


#most uncorrelated below pearson -0.9
most_uncorrelated.neu.sig_09 <- RNA_PROT_cor.neu.sig[RNA_PROT_cor.neu.sig$pearson < -0.9,"Gene",drop = TRUE]
```

```{r}
p1 <- ggplot(subset(RNA_PROT_mean, Gene %in% most_uncorrelated.neu.sig), aes(x = Day, y = norm_exp, group = type, color = type)) +
  geom_point() +
  geom_line() +
  facet_wrap(~Gene, nrow = 2) +
  ylab("scaled expression") +
  theme_classic() +
  ggtitle("Most uncorrelated during neuronal days")

p1b <- ggplot(subset(RNA_PROT_mean, Gene %in% most_uncorrelated.fil.neu), aes(x = Day, y = norm_exp, group = type, color = type)) +
  geom_point() +
  geom_line() +
  facet_wrap(~Gene) +
  ylab("scaled expression") +
  theme_classic() +
  ggtitle("Most uncorrelated during neuronal days (expr. var. > 0.1")

p2 <- ggplot(subset(RNA_PROT_mean, Gene %in% most_uncorrelated.neu.sig_lFC), aes(x = Day, y = norm_exp, group = type, color = type)) +
  geom_point() +
  geom_line() +
  facet_wrap(~Gene) +
  ylab("scaled expression") +
  theme_classic() +
  ggtitle("Most uncorrelated during neuronal days (max. logFC > 2)")

pp1 <- plot_grid(p1,p2,labels = "AUTO",ncol = 2)

ggsave2("./Figure_Output/Most_uncorrelated_neuronal_inclLogFC.png",pp1,width = 11,height = 6,bg = "white")
ggsave2("./Figure_Output/Most_uncorrelated_neuronal.png",p1,width = 10,height = 4,bg = "white")
```


#Create common genes significant vectors
```{r}
common.sig <- intersect(common_genes,sig_all)
common.sig_lFC <- intersect(common_genes,sig_all_lFC)
##no duplicates in those two!

RNA_sig_500.neu <- row.names(se_RNA_red)[order(rowData(se_RNA_red)[["DEA.neuronal"]]$FDR, decreasing = FALSE)[1:500]]
sig.neu.500 <- rowData(se_RNA_red)[RNA_sig_500.neu,]$symbol

RNA_sig_1000.neu <- row.names(se_RNA_red)[order(rowData(se_RNA_red)[["DEA.neuronal"]]$FDR, decreasing = FALSE)[1:1000]]
sig.neu.1000 <- rowData(se_RNA_red)[RNA_sig_1000.neu,]$symbol
```



#Plot side by side
```{r}
# RNA_PROT_merged_long.sig_lFC <- RNA_PROT_merged_long  %>% 
#   dplyr::filter(Gene %in% common.sig_lFC)
# 
# RNA_PROT_scaled <- RNA_PROT_merged_long.sig_lFC  %>%
#   group_by(Gene, type) %>%
#   mutate(scaled_exp = scale(expression))
# 
# ggplot(subset(RNA_PROT_scaled, Gene %in% common.sig_lFC[1:3]), 
#        aes(x = Day, y = scaled_exp, group = interaction(type, replicate), color = replicate, linetype = type)) +
#   geom_point() +
#   geom_line() +
#   facet_wrap(Gene~type, ncol = 2) +
#   theme_bw()
```


##Heatmaps

#Fold change with respect to iPSC stage
```{r}
a <- crossHm(list(RNA = se_RNA_agg, PROTEIN = se_PROT_agg), assayName = "log2FC.neur", features = sig.neu.1000, do.scale = TRUE, top_annotation = "Day", hmcols = heatcol,anno_colors = ancols)

a@ht_list$RNA@top_annotation@anno_list$Day@color_mapping@colors <- ancols$Day
a@ht_list$RNA@top_annotation@anno_list$Day@fun@var_env$color_mapping@colors <- ancols$Day
a@ht_list$PROTEIN@top_annotation@anno_list$Day@color_mapping@colors <- ancols$Day
a@ht_list$PROTEIN@top_annotation@anno_list$Day@fun@var_env$color_mapping@colors <- ancols$Day
a@ht_list$PROTEIN@top_annotation@anno_list$Day@name_param$show <- FALSE
```


```{r}
png("./Figure_Output/RNA_Prot_heat_all_top1000.png", width=11.5, height=6, units="in", res=300)
a
dev.off()
```


```{r}
pdf("./Figure_Output/RNA_Prot_heat_all_top1000.pdf", width=13, height=6)
a
dev.off()
```




## Fold change with respect to first neural stage
```{r}
#Filter neuronal days
se_RNA_agg2 <- se_RNA_agg[,se_RNA_agg$Day != 0]
se_PROT_agg2 <- se_PROT_agg[,se_PROT_agg$Day != 0]

colData(se_RNA_agg2 )$Day <- droplevels(colData(se_RNA_agg2 )$Day)
colData(se_PROT_agg2)$Day <- droplevels(colData(se_PROT_agg2)$Day)
```

```{r}
b <- crossHm(list(RNA = se_RNA_agg2, PROTEIN = se_PROT_agg2), assayName = "log2FC.neur",gaps_at = c("Day"), features = sig.neu.1000, do.scale = TRUE, top_annotation = "Day", hmcols = heatcol,anno_colors = ancols2)

b@ht_list$PROTEIN@top_annotation@anno_list$Day@name_param$show <- FALSE
```


```{r}
png("./Figure_Output/RNA_Prot_heat_neuronal_top1000.png", width=10, height=6, units="in", res=300)
b
dev.off()
```





#Fold change of most uncorrelated RNA/Proteins
```{r}

c <- crossHm(list(RNA = se_RNA_agg2, PROTEIN = se_PROT_agg2), assayName = "log2FC.neur", features = most_uncorrelated.neu.sig_09, do.scale = TRUE, top_annotation = "Day", hmcols = heatcol,anno_colors = ancols2)

c@ht_list$RNA@top_annotation@anno_list$Day@color_mapping@colors <- ancols$Day
c@ht_list$RNA@top_annotation@anno_list$Day@fun@var_env$color_mapping@colors <- ancols$Day
c@ht_list$PROTEIN@top_annotation@anno_list$Day@color_mapping@colors <- ancols$Day
c@ht_list$PROTEIN@top_annotation@anno_list$Day@fun@var_env$color_mapping@colors <- ancols$Day
c@ht_list$PROTEIN@top_annotation@anno_list$Day@name_param$show <- FALSE
```

```{r}
png("./Figure_Output/RNA_Prot_heat_uncorr_neuronal.png", width=10, height=7, units="in", res=300)
c
dev.off()
```






