---
title: "Proteomics_Clustering"
author: "Michael Soutschek"
date: "8 4 2021"
output: html_document
---




```{r}
suppressPackageStartupMessages({
  library(edgeR)
  library(SummarizedExperiment)
  library(SEtools)
  library(ggplot2)
  library(matrixStats)
  library(igraph)
  library(RUVSeq)
  library(pheatmap)
  library(viridis)
  library(plgINS)
  library(stringr)
  library(msigdbr)
})
source("../../../functions/clustering.R")
source("../../../functions/clusterEnrichment.R")
```



# Color Definitions
```{r}
ancols <- list(day=c("0"="#3F3F3F",
                       "9" ="#575757",
                       "15" ="#707070",
                       "21" = "#898989",
                       "27" ="#A1A1A1",
                       "33" ="#BABABA",
                       "40" = "#D3D3D3" 
) )

ancols2 <- list(day=c("0"="#3F3F3F",
                       "9" ="#575757",
                       "15" ="#707070",
                       "21" = "#898989",
                       "27" ="#A1A1A1",
                       "33" ="#BABABA",
                       "40" = "#D3D3D3"),
               cluster=c("1" = "#440154FF",
                         "2" = "#414487FF",
                         "3" = "#2A788EFF",
                         "4" = "#22A884FF",
                         "5" = "#7AD151FF",
                         "6" = "#FDE725FF"
                         ))

heatcol = inferno(256)
```




## Import datasets
```{r}
se <- readRDS("../data/Proteomics/Proteomics.Reference.SE_DEA.rds")
se2 <- se[,se$day != 0]
```



result data.frames
```{r}
res.all <- rowData(se)$DEA.all
res.neu <- rowData(se)$DEA.neuronal

sig.all <- row.names(res.all)[which(res.all$adj.P.Val <0.05)]
sig.neu <- row.names(res.neu)[which(res.neu$adj.P.Val <0.05)]

res.all$maxlfc <- rowMax(abs(as.matrix(res.all[,1:6]))) 
res.neu$maxlfc <- rowMax(abs(as.matrix(res.neu[,1:5]))) 
```


# Clustering all days

## Get preliminary Cluster Info
# Pay attention, takes a very long time
```{r}
## prepare logFC matrix
lfc <- prepLFC4clust(assays(se)$log2FC.all[sig.all,], groups=se$stage, minLFC = 0.5)
## clustering & analysis
png("../../Figures_Omics/Proteomics/Figure_Output/Clustering/Cluster_Info_all-days.png", width=7, height=5, units="in", res=300)
cc <- clusterWrapper(lfc, 2:15, "kmeans")
dev.off()
```


## Get clusters and save object
```{r}
rowData(se)$cluster <- as.factor(cc[["6"]]$cluster[row.names(se)])
saveRDS(se,"../data/Proteomics/Proteomics.SE_DEA.cluster.rds")
```



# Load se object
```{r}
se <- readRDS("../data/Proteomics/Proteomics.SE_DEA.cluster.rds")
```

## smooth plots
```{r}
sm <- getSmoothedMeans(assays(se)$imputed[sig.all,], as.numeric(factor(se$day)), method = "loess", f=0.8)
plotClusterLFCs(sm-rowMeans(sm), rowData(se[sig.all,])$cluster)

png("../../Figures_Omics/Proteomics/Figure_Output/Clustering/DEG_kmeans_all-days.png", width=7, height=5, units="in", res=300)
plotClusterLFCs(sm-rowMeans(sm), rowData(se[sig.all,])$cluster)
dev.off()
```


# Cluster Enrichment Analysis
```{r}
all.clus <- split(rowData(se)$GeneName,rowData(se)$cluster)
all.clus.un <- lapply(all.clus, unique)

m <- msigdbr(db_species = "HS",species = "Homo sapiens")
m <- m[which(m$gs_collection %in% "C5"), ]
m$name2 <- paste0(m$gs_collection, ":", m$gs_subcollection, ":", m$gs_name)
gsets <- split(m$gene_symbol, m$name2)

gsets_cl <- gsets[!grepl("C5:HPO",names(gsets))]
names(gsets_cl) <- gsub("C5:GO:BP:","",names(gsets_cl))
names(gsets_cl) <- gsub("C5:GO:CC:","",names(gsets_cl))
names(gsets_cl) <- gsub("C5:GO:MF:","",names(gsets_cl))
dgo.all <- clusterEnrichment(all.clus.un, gsets_cl)
```

```{r}
a <- plotClusterEnrichment(dgo.all,color = heatcol)
png("../../Figures_Omics/Proteomics/Figure_Output/Clustering/DEG_kmeans_ClusterEnrichment_all-days.png", width=12, height=10, units="in", res=300)
draw(a,heatmap_legend_side="left")
dev.off()
```





# Clustering neuronal days object

## Get preliminary Cluster Info
```{r}
## prepare logFC matrix
lfc2 <- prepLFC4clust(assays(se2)$log2FC.neur[sig.neu,], groups=se2$stage, minLFC = 0.5)
lfc2 <- lfc2[,-1]
## clustering & analysis
png("../../Figures_Omics/Proteomics/Figure_Output/Clustering/Cluster_Info_Neur-days.png", width=7, height=5, units="in", res=300)
cc2 <- clusterWrapper(lfc2, 2:15, "kmeans")
dev.off()
```

## Get clusters and save object
```{r}
rowData(se2)$cluster <- as.factor(cc2[["6"]]$cluster[row.names(se2)])
saveRDS(se2,"../data/Proteomics/Proteomics.SE_DEA.SE2.neur_days.cluster.rds")
```

# Load se2 object
```{r}
se2 <- readRDS("../data/Proteomics/Proteomics.SE_DEA.SE2.neur_days.cluster.rds")
```

## smooth plots
```{r}
sm2 <- getSmoothedMeans(assays(se2)$imputed[sig.neu,], as.numeric(factor(se2$day)), method = "loess", f=0.8)
plotClusterLFCs(sm2-rowMeans(sm2), rowData(se2[sig.neu,])$cluster)

png("../../Figures_Omics/Proteomics/Figure_Output/Clustering/DEG_kmeans_neur-days.png", width=8, height=5, units="in", res=300)
plotClusterLFCs(sm2-rowMeans(sm2), rowData(se2[sig.neu,])$cluster) + theme_classic() +
  scale_x_discrete(
    breaks = 1:6,
    labels = c("Day9", "Day15", "Day21", "Day27", "Day33", "Day40"))
dev.off()
```



# Cluster Enrichment Analysis
```{r}
neu.clus <- split(rowData(se2)$GeneName,rowData(se2)$cluster)
neu.clus.un <- lapply(neu.clus, unique)
m <- msigdbr(db_species = "HS",species = "Homo sapiens")
m <- m[which(m$gs_collection %in% "C5"), ]
m$name2 <- paste0(m$gs_collection, ":", m$gs_subcollection, ":", m$gs_name)
gsets <- split(m$gene_symbol, m$name2)

gsets_cl <- gsets[!grepl("C5:HPO",names(gsets))]
names(gsets_cl) <- gsub("C5:GO:BP:","",names(gsets_cl))
names(gsets_cl) <- gsub("C5:GO:CC:","",names(gsets_cl))
names(gsets_cl) <- gsub("C5:GO:MF:","",names(gsets_cl))
dgo.neu <- clusterEnrichment(neu.clus.un, gsets_cl)
```

```{r}
#Plot a clean version
dgo.neu2 <- dgo.neu
row.names(dgo.neu2) <- gsub("GOBP_","",row.names(dgo.neu2))
row.names(dgo.neu2) <- gsub("GOCC_","",row.names(dgo.neu2))
row.names(dgo.neu2) <- gsub("GOMF_","",row.names(dgo.neu2))

b <- plotClusterEnrichment(dgo.neu2,color = heatcol,trim.names = 35L,row_names_max_width = unit(9, "cm"),)

png("../../Figures_Omics/Proteomics/Figure_Output/Clustering/DEG_kmeans_ClusterEnrichment_neur-days.png", width=12, height=9, units="in", res=300)
draw(b,heatmap_legend_side="left")
dev.off()
```




