---
title: "TopGo Compartment"
author: "Michael"
date: "02.07.2020"
output: html_document
---

```{r}
library(topGO)
library(org.Hs.eg.db)
library(openxlsx)
```

#script executed with org.Hs.eg.db (version 3.12)

#load experiment
```{r}
se <- readRDS("../../data/SE.dea.Gene.salmon.raw.rds")

dea1229 <- rowData(se)[["DEA.ind.Ctrl-MiR"]] 
dea1229$Ensembl <- sapply(strsplit(row.names(dea1229),split = "\\."),"[",1)
dea1229$Symbol <- sapply(strsplit(row.names(dea1229),split = "\\."),"[",2)
```



#Create the Vectors

Dend: All significants vs the genes from the specific background
```{r}
GeneUniverse <- as.character(dea1229$Symbol)
GenesOfInterest <- as.character(dea1229[dea1229$FDR < 0.05 & dea1229$logFC > 0,"Symbol"])
GeneList <- factor(as.integer(GeneUniverse %in% GenesOfInterest))
names(GeneList) <- GeneUniverse
head(GeneList)
```

#Preparing the TopGO data object. Pay attention on the ID value, that depends on what you've
#inserted before in the GeneList. We do the analysis for all three ontologies

```{r}
tgdBP <- new( "topGOdata", ontology="BP", allGenes = GeneList, nodeSize=5,
            annot=annFUN.org, mapping="org.Hs.eg.db", ID = "symbol" )
tgdMF <- new( "topGOdata", ontology="MF", allGenes = GeneList, nodeSize=5,
              annot=annFUN.org, mapping="org.Hs.eg.db", ID = "symbol" )
tgdCC <- new( "topGOdata", ontology="CC", allGenes = GeneList, nodeSize=5,
              annot=annFUN.org, mapping="org.Hs.eg.db", ID = "symbol" )
```

#Afterwards one can run several tests, depending on algorithm and statistics
```{r}
resultTopGO.elim_BP <- runTest(tgdBP, algorithm = "elim", statistic = "Fisher" )
resultTopGO.classic_BP <- runTest(tgdBP, algorithm = "classic", statistic = "Fisher" )

resultTopGO.elim_MF <- runTest(tgdMF, algorithm = "elim", statistic = "Fisher" )
resultTopGO.classic_MF <- runTest(tgdMF, algorithm = "classic", statistic = "Fisher" )

resultTopGO.elim_CC <- runTest(tgdCC, algorithm = "elim", statistic = "Fisher" )
resultTopGO.classic_CC <- runTest(tgdCC, algorithm = "classic", statistic = "Fisher" )
```

#And one can display the results in a GeneTable
```{r}
tabBPF <- GenTable( tgdBP, Fisher.elim = resultTopGO.elim_BP, 
                      Fisher.classic = resultTopGO.classic_BP,
                      orderBy = "Fisher.classic" , topNodes = 200)
tabBPE <- GenTable( tgdBP, Fisher.elim = resultTopGO.elim_BP, 
                   Fisher.classic = resultTopGO.classic_BP,
                   orderBy = "Fisher.elim" , topNodes = 200)


tabMFF <- GenTable( tgdMF, Fisher.elim = resultTopGO.elim_MF, 
                   Fisher.classic = resultTopGO.classic_MF,
                   orderBy = "Fisher.classic" , topNodes = 200)
tabMFE <- GenTable( tgdMF, Fisher.elim = resultTopGO.elim_MF, 
                    Fisher.classic = resultTopGO.classic_MF,
                    orderBy = "Fisher.elim" , topNodes = 200)

tabCCF <- GenTable( tgdCC, Fisher.elim = resultTopGO.elim_CC, 
                   Fisher.classic = resultTopGO.classic_CC,
                   orderBy = "Fisher.classic" , topNodes = 200)
tabCCE <- GenTable( tgdCC, Fisher.elim = resultTopGO.elim_CC, 
                   Fisher.classic = resultTopGO.classic_CC,
                   orderBy = "Fisher.elim" , topNodes = 200)
```



# Do also the FDR correction in R and add to the GenTable data.frame as a sepperate object
```{r}
tabBPF$Classic_Fishe_BH_Corr = p.adjust(tabBPF$Fisher.classic, method = "BH")
tabBPE$Classic_Fishe_BH_Corr = p.adjust(tabBPE$Fisher.classic, method = "BH")
tabMFF$Classic_Fishe_BH_Corr = p.adjust(tabMFF$Fisher.classic, method = "BH")
tabMFE$Classic_Fishe_BH_Corr = p.adjust(tabMFE$Fisher.classic, method = "BH")
tabCCF$Classic_Fishe_BH_Corr = p.adjust(tabCCF$Fisher.classic, method = "BH")
tabCCE$Classic_Fishe_BH_Corr = p.adjust(tabCCE$Fisher.classic, method = "BH")
```



## Adding Fold Enrichment
```{r}
tabBPF$Fold_Enrichment = tabBPF$Significant / tabBPF$Expected
tabBPE$Fold_Enrichment = tabBPE$Significant / tabBPE$Expected
tabMFF$Fold_Enrichment = tabMFF$Significant / tabMFF$Expected
tabMFE$Fold_Enrichment = tabMFE$Significant / tabMFE$Expected
tabCCF$Fold_Enrichment = tabCCF$Significant / tabCCF$Expected
tabCCE$Fold_Enrichment = tabCCE$Significant / tabCCE$Expected
```



## Sorting the columns
```{r}
tabBPF <- tabBPF[c(1,2,3,4,5,10,6,7,8,9)]
tabBPE <- tabBPE[c(1,2,3,4,5,10,6,7,8,9)]
tabMFF <- tabMFF[c(1,2,3,4,5,10,6,7,8,9)]
tabMFE <- tabMFE[c(1,2,3,4,5,10,6,7,8,9)]
tabCCF <- tabCCF[c(1,2,3,4,5,10,6,7,8,9)]
tabCCE <- tabCCE[c(1,2,3,4,5,10,6,7,8,9)]
```



## And do the GeneList annotation in R
# First get a vector of the GenesOfInterest ("significant Genes"), then annotate
#them to the Table (for all of the tables)

#For Biological Processes
```{r}
SigGenesBP <- sigGenes(tgdBP)

tabBPF$Sig_Genes <- sapply(tabBPF$GO.ID, function(x)
  {
  Sig_Genes <- genesInTerm(tgdBP, x)
  Sig_Genes[[1]][Sig_Genes[[1]] %in% SigGenesBP]
})

tabBPE$Sig_Genes <- sapply(tabBPE$GO.ID, function(x)
{
  Sig_Genes <- genesInTerm(tgdBP, x)
  Sig_Genes[[1]][Sig_Genes[[1]] %in% SigGenesBP]
})

```


#For Molecular Function
```{r}
SigGenesMF <- sigGenes(tgdMF)

tabMFF$Sig_Genes <- sapply(tabMFF$GO.ID, function(x)
{
  Sig_Genes <- genesInTerm(tgdMF, x)
  Sig_Genes[[1]][Sig_Genes[[1]] %in% SigGenesMF]
})

tabMFE$Sig_Genes <- sapply(tabMFE$GO.ID, function(x)
{
  Sig_Genes <- genesInTerm(tgdMF, x)
  Sig_Genes[[1]][Sig_Genes[[1]] %in% SigGenesMF]
})

```


#For Cellular Components
```{r}
SigGenesCC <- sigGenes(tgdCC)

tabCCF$Sig_Genes <- sapply(tabCCF$GO.ID, function(x)
{
  Sig_Genes <- genesInTerm(tgdCC, x)
  Sig_Genes[[1]][Sig_Genes[[1]] %in% SigGenesCC]
})

tabCCE$Sig_Genes <- sapply(tabCCE$GO.ID, function(x)
{
  Sig_Genes <- genesInTerm(tgdCC, x)
  Sig_Genes[[1]][Sig_Genes[[1]] %in% SigGenesCC]
})
```


#saveRDS
```{r}
saveRDS(tabBPE, file = "./GOdata/GO_1229_BP_sig_005_up_symbol_elim.rds")
saveRDS(tabMFE, file = "./GOdata/GO_1229_MF_sig_005_up_symbol_elim.rds")
saveRDS(tabCCE, file = "./GOdata/GO_1229_CC_sig_005_up_symbol_elim.rds")
```

```{r}
#rm(list=ls())
```


```{r}
# rm(ExcelFile,resultTopGO.classic_BP,resultTopGO.classic_CC,resultTopGO.classic_MF,resultTopGO.elim_BP,resultTopGO.elim_CC,resultTopGO.elim_MF,
#    tabBPE,tabBPF,tabCCE,tabCCF,tabMFE,tabMFF,
#    tgdBP,tgdCC,tgdMF,GeneList,GenesOfInterest,GeneUniverse,
#    SigGenesBP,SigGenesCC,SigGenesMF)
```

               


